Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to MIR downstream tasks

### 2023-06-08 abstract/abstract.tex
Updates from Overleaf

In collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), we used self-supervised contrastive learning to uncover the underlying structure of Western tonal music. We applied deep convolutional neural networks and a triplet loss function, identifying abstract and semantic musical elements without relying on their sonic qualities. This way, we replaced traditional acoustic features with deep audio embeddings, paving the way for high-level, sound-agnostic, and content-sensitive music representation.

Our cognitively-based approach for learning embeddings focuses on using full-resolution data and preserving high-level musical information. Drawing upon our domain expertise, we developed robust transformations to encode heuristic musical concepts that should remain constant. This novel approach aims to reconcile music and machine learning, enhancing machine listening models' efficacy. Preliminary results suggest that our musically-informed technique has significant potential for boundary detection tasks and possibly nearly all MIR downstream tasks that are not purely sonic-based. While it didn't outperform state-of-the-art results, it showed very competitive results and room for improvement. Consequently, music-motivated audio embeddings appear promising and potentially adaptable to other tasks constrained by data scarcity.

>>>>

In collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), we used self-supervised contrastive learning to uncover the underlying structure of Western tonal music by learning deep audio features to improve unsupervised music boundary detection. We applied deep convolutional neural networks and a triplet loss function to identify abstract and semantic musical elements without relying on their sonic qualities. This way, we replaced traditional acoustic features with deep audio embeddings, paving the way for high-level, sound-agnostic, and content-sensitive music representation for boundary detection.

Our cognitively-based approach for learning embeddings focuses on using full-resolution data and preserving high-level musical information. Drawing upon our domain expertise, we developed robust transformations to encode heuristic musical concepts that should remain constant. This novel approach aims to reconcile music and machine learning, enhancing machine listening models' efficacy.

Preliminary results suggest that our musically-informed technique has significant potential for boundary detection tasks and possibly nearly all MIR downstream tasks that are not purely sonic-based. While it didn't outperform state-of-the-art results, it showed very competitive results and room for improvement. Consequently, music-motivated audio embeddings appear promising and potentially adaptable to other tasks constrained by data scarcity.

### 2023-06-08 future work/future work.tex
Updates from Overleaf

Listening evaluation
2D or 3D latent space visualization as per \ref{fig:manifold}

>>>>

Visual and listening evaluation: 2D or 3D latent space visualization as per \ref{fig:manifold}. Arranging the embedding space in a visual display to evaluate the clustering of this sophisticated musical content to see to what extent they consider sonic attributes.

### 2023-06-08 introduction/Subsections/Objectives.tex
Updates from Overleaf

This study explores music's high-level tonal structure by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn and analyze music embeddings applicable to subsequent MIR tasks.

The study aims to develop a unified embedding, $e(M)$, for a music piece $M$ that integrates high-level elements like melody contour, harmony relationships, and other comprehensive components, arranging the embedding space to emphasize clustering of these sophisticated musical content without any consideration for sonic quality attributes. Utilizing a blend of Music Information Retrieval (MIR) computational methodologies and musicological perspectives, the research will explore musical compositions from various styles and genres.

>>>>

Can we learn deep audio embeddings from reordering and scrambling sequences of musical information and adding audio effects to improve unsupervised music boundary detection?

This study explores music's high-level tonal structure by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn and analyze music embeddings applicable to subsequent MIR tasks, in our particular case, music boundary detection.

The study aims to develop and learn a unified numerical understanding (or embedding \footnote{A learned representation or embedding is a numerical output -usually a fixed-size vector- produced by a machine-learning model. Good representations should be versatile across various tasks and require limited supervision \cite{Turian2022HEAR:Representations}.}) for a music piece that integrates high-level elements like melody contour, harmony relationships, and other comprehensive components such as the relationships between elements unfolding through time.

The research will utilize MIR computational methodologies and musicological perspectives to explore musical compositions from various styles and genres.

### 2023-06-08 methods/subsections/implementation.tex
Updates from Overleaf

Conventional transformations, such as log-scaled Mel spectrograms or Mel Frequency Cepstral Coefficients, capture key features but may obscure some information inherent in raw audio data.

Raw audio data offers unique advantages. Primarily, it ensures the retention of the original signal, which could reveal novel findings, and facilitates the direct extraction of features through advanced deep learning models \cite{learning}. However, it is not without challenges: high sample rates result in high dimensionality, demanding substantial computational resources. Additionally, raw audio is noise-sensitive, complicating the extraction process. Features derived from raw audio also often lack the interpretability of traditional ones \cite{Schindler2020DeepTutorial}. Hence, the choice of input data depends on the specific task, available computational resources, and the balance between data preservation and computational efficiency.

Time-domain processing, a feature of our experiments, naturally handles temporal patterns and sequences in data, thereby avoiding windowing artifacts. Although Mel-spectrogram-based methods are effective for various audio-related machine learning tasks, their limitations lie in representing perceptual similarity. Spectrograms capture frequency distribution over time. Yet, the complexity of human auditory perception—encompassing temporal patterns, phase relationships between frequencies, and higher-level musical structures—means that musically similar sounds can have distinct spectrograms. This discrepancy implies that using spectrogram distance alone for measuring high-level music similarity may not always align with human perceptions \cite{Kim2020OneStrategies}\cite{Mesostructures2023}.

@@

The positive image in our study must preserve its intelligible content when subjected to transformations, regardless of alterations in sonic qualities. Maintaining the temporal structure and meaningfulness of the content allows it to present musical elements identical or strikingly similar to the original track.

While we have experimented with helpful audio augmentation tools such as \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0} and \cite{Kharitonov2020DataDomain}, the specific requirements of our experiments necessitated the development of our own transformation chain using \textit{torchaudio's} \cite{Yang2021TorchAudio:Processing} implementation of \textit{SoX} \cite{sox}: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

In this analysis, various parameters represent different effects on the signal. We use $g \in [-12, 0]$ for gain, $\Delta p \in [-1200, 1200]$ for pitch shift, $h_R[n]$ for the reverb impulse response, $h_C[n]$ for the chorus impulse response, $d \in [0, 30]$ for overdrive, $\alpha \in [0.9, 1.1]$ for speed change, $\beta \in [0.9, 1.1]$ for stretch factor, $f_m \in [0.1, 100]$ for tremolo's modulation frequency, and $d_m \in [1, 101]$ for tremolo's depth. Each effect is applied aggressively but does not disrupt the high-level musical content.

\textbf{Amplitude effects}, like gain, modify the signal's overall amplitude with a constant factor $g$.

\textbf{Time-domain effects} such as speed change and stretch alter the signal's playback speed and duration without changing its pitch, respectively. The speed and stretch factors are $\alpha$ and $\beta$.

\textbf{Frequency-domain effects} adjust the frequency content or pitch of the signal, as in pitch shifting, which alters the pitch by $\Delta p$.

\textbf{Nonlinear effects} such as overdrive add harmonic distortion using the parameter $d$.

\textbf{Modulation effects}, including chorus and tremolo, use a control signal or low-frequency oscillator. The specified ranges determine the chorus parameters, and the tremolo's amplitude modulation frequency and depth are $f_m$ and $d_m$.

\textbf{Reverberation effects} simulate physical space's acoustic reflections and reverberations, like reverb that applies an impulse response $h_R[n]$. Finally, \textbf{Noise effects} add a noise signal $N[n]$ with a signal-to-noise ratio (SNR) in the range $[12, 100]$.

@@

The minimum duration $t_{min}$ is set to 0.05 seconds, and the maximum duration $t_{max}$ is set to 1 second. This range is chosen thoughtfully to strike a balance between the two factors. On the one hand, it is above the just noticeable difference (JND) threshold, the smallest change in a stimulus that can be perceived. On the other hand, it is short enough to maintain a reasonably-sized window to avoid discernible musical content \cite{Fastl2007Just-NoticeableChanges}.

@@

\item Concatenate the shuffled audio chunks to generate the negative signal:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

>>>>

Features derived from raw audio often lack the interpretability of traditional ones \cite{Schindler2020DeepTutorial}. Hence, the choice of input data depends on the specific task, available computational resources, and the balance between data preservation and computational efficiency.

While previous studies have used CNNs and various features as input like Mel-Scaled Log-magnitude Spectograms (MLS), Self-Similarity Matrices (SSM), and Self-Similarity Lag Matrices (SSLM) as inputs \cite{Hernandez-Olivan2021MusicFeatures}, raw audio data offers unique advantages: it ensures the retention of the original signal, which could reveal novel findings, and facilitates the direct extraction of features through advanced deep learning models \cite{learning}. However, it is not without challenges: high sample rates result in high dimensionality, demanding substantial computational resources. Additionally, raw audio is noise-sensitive, complicating the extraction process.

Time-domain processing naturally handles temporal patterns and sequences in data, thereby avoiding windowing artifacts. Although Mel-spectrogram-based methods are effective for various audio-related machine learning tasks, their limitations lie in representing perceptual similarity. Spectrograms capture frequency distribution over time. Yet, the complexity of human auditory perception—encompassing temporal patterns, phase relationships between frequencies, and higher-level musical structures—means that musically similar sounds can have distinct spectrograms. This discrepancy implies that using spectrogram distance alone for measuring high-level music similarity may not always align with human perceptions \cite{Kim2020OneStrategies, Mesostructures2023}.

@@

The positive image in every triplet of input data must preserve its intelligible content when subjected to transformations, regardless of alterations in sonic qualities. Maintaining the temporal structure and meaningfulness of the content allows it to present musical elements identical or strikingly similar to the original track.

While we have experimented with helpful audio augmentation tools \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0, Kharitonov2020DataDomain}, the specific requirements of our experiments necessitated the development of our transformation chain using \textit{torchaudio}'s \cite{Yang2021TorchAudio:Processing} implementation of \textit{SoX} \cite{sox}: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

\textbf{Amplitude effects}, like gain, modify the signal's amplitude with a constant factor $g$. We use $g \in [-12, 0]$.

\textbf{Time-domain effects} such as speed change and stretch alter the signal's playback speed and duration without changing its pitch, respectively. The speed and stretch factors are $\alpha \in [0.9, 1.1]$ for speed change and $\beta \in [0.9, 1.1]$ for the stretch factor.

\textbf{Frequency-domain effects} adjust the frequency content or pitch of the signal, as in pitch shifting, which alters the pitch by $\Delta p \in [-1200, 1200]$.

\textbf{Nonlinear effects} such as overdrive add harmonic distortion using the parameter $d \in [0, 30]$ for overdrive

\textbf{Modulation effects}, including chorus and tremolo, use a control signal or low-frequency oscillator. The specified ranges determine the chorus parameters and the tremolo's amplitude modulation frequency and depth $f_m \in [0.1, 100]$ for the tremolo's modulation frequency and $d_m \in [1, 101]$ for the tremolo's depth

\textbf{Reverberation effects} simulate physical space's acoustic reflections and reverberations, like reverb that applies an impulse response $h_R[n]$.

Finally, \textbf{Noise effects} add a noise signal $N[n]$ with a signal-to-noise ratio (SNR) in the range $[12, 100]$.

@@

The minimum duration $t_{min}$ is set to 0.05 seconds, and the maximum duration $t_{max}$ is set to 1 second. This range is chosen thoughtfully to strike a balance between two factors: on the one hand, it is above the just noticeable difference (JND) threshold, the smallest change in a stimulus that can be perceived. On the other hand, it is short enough to maintain a reasonably-sized window to avoid discernible musical content \cite{Fastl2007Just-NoticeableChanges}.

@@

\item We finally concatenate the shuffled audio chunks to generate the negative signal:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

### 2023-06-08 results/subsections/tables.tex
Updates from Overleaf

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Year, Authors, Ref}} & \thead{\centering\textbf{Algorithm}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
2012, Kaiser et al., [27]& KSP2 & SSM & Novelty measure  & 0.286 \\
\addlinespace
2013, McFee \& Ellis, [20] & MP2 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
2014, Nieto \& Bello, [28] & NB1 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
2015, Cannam et al., [29] & CC1 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
2016, Nieto, [30] & ON2 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
2017, Cannam et al., [29] & CC1 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
2007, Turnbull et al., [19] & - & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
2011, Sargent et al., [34] & - & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
2014, Ullrich et. al, [22] & - & MLS & CNN  & 0.465 \\
\addlinespace
2015, Grill \& Schlüter, [4] & - & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
2015, Grill \& Schlüter, [5] & - & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
2017, Hadria \& Peeters, [35] & - & MLS, SSLMs & CNN  & 0.291 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State of the art table.]{Your Caption.}
\label{tab:my_label}
\end{table}

>>>>

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Authors [Ref], Year}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{GrillMUSICANNOTATIONS}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State of the art table.]{Your Caption.}
\label{tab:my_label}
\end{table}
